Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-09-07 Final_Manuscript.tex
re-knitted word and pdf

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for woman within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between such valuable phenomena and their effects on authenticity within
interactions between female pairs. Woman were brought into the
laboratory and interacted in same-sex dyads. Dyadic analysis was
utilized to detect whether partners' objectification of each other
affected state self-objectification, and the resulting feelings of
comfort and authenticity during the interaction. After the interaction,
participants completed a questionnaire which measured many constructs
including cognitive performance, career aspirations, and relationship
agency. Results revealed no significant relationship between
self-objectification and authenticity. Further, although there were
significantly negative effects on career aspirations and relationship
agency resulting from a lack of relationship authenticity, there was no
evidence that this is due to feelings of sexual objectification. The
significant partner effect of objectification on actor
self-objectification suggests that women being objectified by other
women still results in feelings of self-objectification, and such
research has powerful implications for the ways that women interact in
both sexual and non-sexual settings. AUTHOR NOTE: mention Clark trigram
coders Hannah et al.

@@

Roberts and Fredrickson (1997) Objectification Theory, suggests that
women are objectified in interpersonal encounters, and that the effect
might be strongest from men, but possibly also from women. Psychological
researchers studying the sexual objectification of women, and the
resulting self-objectification that follows, have in recent years
explored the interpersonal process of the objectification -
self-objectification link (CITE INTERPERSONAL PROCESS LINKS LITERATURE).
For example, in a face-to-face mixed-gender dyadic study, Garcia,
Earnshaw, and Quinn (2016) found that men's reported objectification of
their female interaction partner during an interaction was associated
with increased self-objectification reported by their partner. Gervais,
Sáez, Riemer, and Klein (2020) has reviewed the literature on
interpersonal self-objectification and has organized the the process in
a theoretical model called the SIMO, but this model is only relevant for
mixed-gender encounters. Although there is ample evidence that women
objectify other women (Loughnan et al., 2015a; Puvia \& Vaes, 2013),
there is currently no studies investigating the process of interpersonal
objectification same-gender interactions. Does objectification by a
female interaction partner have the same downstream negative
consequences as being objectified by a male interaction partner? For
example, perhaps women could be amplifying each others'
self-objectification as they get ready to go out for the night,
\emph{before} the theoretical encounters with men outlined in Gervais et
al. (2020) occur.

@@

Self-objectification is a multidimensional process that accounts for the
cognitive mechanism that translates experiences of sexualization at the
cultural level (Loughnan et al., 2015a) to psychological (e.g., anxiety,
self-esteem, authenticity, motivational states) and behavioral
(cognitive performance, body monitoring) features of mental health and
well-being at the individual level (Calogero, Tantleff-Dunn, \&
Thompson, 2011; Moradi \& Huang, 2008). Calogero et al. (2011) proposes
that the construct of self-objectification can be conceptualized as a
learned trait. Furthermore, it can also be elicited momentarily, through
the media, for example, with sexualized images in movies and magazines
(Morry \& Staska, 2001), which can lead to a state of
self-objectification (Bernard, Legrand, \& Klein, 2018; Calogero et al.,
2011; Moradi \& Huang, 2008). Being objectified by another person and
possessing trait-level self-objectification (TSO) may interact to
influence experiences of feeling like a body, or state
self-objectification (SSO) (B. L. Fredrickson, Roberts, Noll, Quinn, \&
Twenge, 1998, Garcia et al. (2016)). Does this habit
self-objectification lead women to also objectify other women, possibly
to compare oneself?

@@

Past research has focused on men's objectifying behaviors, with little
attention to the role that women may also play in enacting
objectification of women. Hill and Fischer (2008) assessed women's
experiences of objectification from men separately from women's
experiences of objectification from women and found that women may be
socialized not only to see themselves as objects but perhaps to see
other women as objects as well (THEY DIDN'T DETERMINE THIS IN PERSON).
Some studies have found that women objectify other women to a greater
extent than they objectify men (Loughnan et al., 2015a). Women objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but we know next to nothing about interpersonal sexual
objectification among women and what the immediate effects might be.
Further, do women objectify non-sexualized women they are interacting
with? What might be different for same-gender interactions? Harsey and
Zurbriggen (2020) found that self-objectification was related to the
objectification of women to a similar degree for men and women
participants, although men scored high than women on a measure of belief
that objectification of women is natural for men. ASK RANDI-WE KNOW
OBJECTIFICATION HAS MORE ADVERSE CONSEQUENCES FOR WOMEN THAN MEN
(Gervais, Vescio, \& Allen, 2011; Moradi \& Huang, 2008; Saguy, Quinn,
Dovidio, \& Pratto, 2010)\ldots{}MENTION THAT\ldots{}HOWEVER DOES WOMEN
OBJECT WOMEN TO A GREATER EXTENT THAN MEN MATTER MORE?\ldots{}STILL
TRYING TO FIGURE THAT OUT? ADDITIONALLY Strelan and Hargreaves (2005)
FOUND THAT THE MORE WOMEN SELF-OBJECTIFY, THE MORE THEY OBJECTIFY OTHER
WOMEN.

@@

SIMO model and literature review (Gervais et al., 2020). Studies have
shown that within social encounters women are gazed at more than men
(Briton \& Hall, 1995), often times feel \enquote{looked at} within
interpersonal interactions (Argyle \& Williams, 1969; Miles-McLean et
al., 2015), and will more than likely internalize the objectifying gaze
on physical self (Puvia \& Vaes, 2013; Young, 1979). Moreover, perhaps
the most adverse effect of objectifying treatment is that it effectively
socializes girls and women to treat themselves as objects to be looked
at and evaluated, an effect termed self-objectification (Bartky, 1990;
Berger, Cohen, \& Zelditch Jr, 1972; B. L. Fredrickson et al., 1998). We
don't know about women-on-women interpersonal objectification. What
might be different?

Self-objectification has been found to occur after a mere relationship
prime among women (Sanchez \& Broccoli, 2008) because in our culture
women need to look attractive to obtain and maintain successful
relationship, thus, the ``college relationships condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. Among heterosexual male and female college students, self-
objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, and women can sometimes
have positive outcomes from appearance valuation within the context of
an established romantic relationship and they experience less SSO after
appearance comments within romantic relationships (Meltzer, 2020), less
is known about first dates. But this literature is also only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) {[}CITE PWQ paper I
reviewed, Yoder editor{]}. Empirical evidence reveals that
objectification manifests through inauthenticity in romantic
relationships (Brunell et al., 2010), adverse attitudes in regard to
career aspirations, and a decrease in concentration and impairment in
female cognitive performance (Kahalon, Shnabel, \& Becker, 2018; D. M.
Quinn, Chaudoir, \& Kallen, 2011). Both romantic relationships and
relationships in general have been studies in the context of
objectification, but neither of these literatures have investigated
specifically same-gender relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) and within the self in any
moment (trying on a swimsuit) has negative consequences, but what about
in the context of interaction with other women? Is the
self-objectification experience in mixed-gender interactions the same as
that experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
would suggest no (Gay \& Castano, 2010; Gervais, Vescio, \& Allen, 2011;
Saguy, Quinn, F Dovidio, \& Pratto, 2010, Calogero (2004); Yilmaz \&
Bozo, 2019). Women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013), but when they do it, does
it lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

There is some evidence that reduced authenticity is a consequence of
self-objectification in the moment {[}Garcia et al. (2016); again CITE
PWQ paper I reviewed, Yoder editor{]}. This link has been justified by
the literature on stigmatized-stigmatizer interactions (CITE LITERATURE
DIDN'T COME ACROSS THIS), viewing the experience of being objectified in
an interaction as an identity threat situation (MAYBE CITE MILES-MCLEAN
PAPER \enquote{STOP LOOKING AT ME!}). When a woman is objectified by a
man, and then experiences self-objectification, this intergroup
encounter might trigger identitiy threat, but is the same negative
consequence present when a woman is objectified by another woman?
woman-woman interpersonal objectification processes might diverge from
this process. One the one hand, she may not be having the same negative
consequences the cascade from situations that threated group-based
identity (Deaux \& Major, 1987; Dovidio, Hebl, Richeson, \& Shelton,
2006; M. R. Hebl \& Dovidio, 2005), but feeling like a body, rather than
a full human, in an interaction may be enough to reduce feelings of
authenticity and social competence, regardless of the gender of the
objectifyer {[}Tolman, Impett, Tracy, and Michael (2006); PWQ paper I
reviewed, Yoder editor{]}. I DONT KNOW WHERE THIS PAPER IS? ASK RANDI
ABOUT IDENTITY\_THREAT PAPERS

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-sex female
interpersonal interactions. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. We expected to
replicate the results found in Garcia et al. (2016). We predicted that
being objectified by one's interaction partner would lead to
self-objectification, which in turn would lead to feelings of
inauthenticity, then reduced feelings of agency in romantic
relationships, reduced career aspiration, and reduced cognitive
performance. Specifically, we expected to find a positive relationship
between other-objectification by one's partner and state
self-objectification. We also expect to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

Puvia and Vaes (2013) would alternatively predict that women's tendency
to self-objectify (TSO) leads them to objectify other women
(\ldots{}well, dehumanize a sexualized woman, but also Strelan and
Hargreaves (2005) would predict the TSO-\textgreater{}SOO link and
Harsey and Zurbriggen (2020) too), and this is mediated by the woman's
own state self-objectification (SSO). So we also tested this as a
possible alternative model.

@@

The procedure used was identical to that in Garcia et al. (2016), except
for the instructions that the participants were given. In brief, that
methodology is that each participant arrived at the laboratory and were
then led into separate cubicles to prevent any communication between the
participants before the interaction. In addition, each participant was
screened for prior acquaintance to confirm that they had not met prior
to the study. They were asked to sign the consent form to participate,
and the study was described as follows: \enquote{This is a study looking
at how students form different types of relationships at college.} A
prompt on the computer screen told the participants that they were
assigned to the \enquote{College Relationships} condition and gave the
following instructions:

\begin{quote}
There are many types of relationships people form in college. During the
interaction, please think about your partner's potential as a romantic
partner. Even if they are not the gender you are attracted to, you can
still judge their potential as a romantic partner. After the interaction
you will be asked to evaluate how dateable your partner is. In other
words, we would like to know if you think someone would date your
interaction partner. Also, your interaction partner will be evaluating
you in the same manner.
\end{quote}

The decision was made to ask even heterosexual women to judge their
women partners as potential romantic partners. We felt that this prompt
would keep the study closest to a replication of the previous Garcia et
al. (2016) version of the study. Past research has found that women,
largely due to an increase self-objectifying state, are indeed able to
evaluate other women's potential as romantic partners--indeed, women may
be unfortunately used to thinking this way about themselves, and we
suspect other women.

Two participants were then brought into a larger interaction room where
they sat on stools prearranged to be 36 inches apart. The experimenter
instructed the participants to \enquote{get to know each other} for 10
minutes and then left the room. After 10 minutes, the experimenter came
back into the room and stopped the interaction. The participants then
went back to their individual cubicles and completed a set of
post-interaction measures. Participants were then thanked for their
participation and debriefed Garcia et al. (2016). The full methodology
used is found in Garcia et al. (2016)'s study.

@@

Due these similarities across samples in regard to correlation patterns
between significant variables within this study, the two datasets were
combined. These participants were mostly first-year college students,
with an average age of 18.85 (SD = 1.04). The sample was 48.44\%
White/European American, 9.38\% Black/African-American, 28.12\%
Asian/Pacific Islander, 9.38\% Latinx, and 4.69\% mixed-race. There were
8 White/White pairs and 4 same race racial minority pairs, for a total
of 12 same-race pairs. The remaining 20 were mixed race pairs, of which
15 were White/racial minority pairings and 5 were cross-racial minority
group pairs. 64.06\% of the sample identified as heterosexual, and 25\%
identified as gay, lesbian or bisexual.

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner Garcia et al. (2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were to be rated on
a scale from 1 (not at all) to 7 (constantly). Objectification was
measured by getting the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

@@

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much alike (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{`Do you think your partner
was authentic during your interaction?}' and \enquote{`Were you
authentic during your interaction?}' These questions were ranked on a
scale from 1 (not authentic at all). These were combined to form the
authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) = 0.91
for Sample 2).

@@

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{apaTables} (Version 2.0.5; Stanley, 2018), \emph{devtools}
(Version 1.13.5; Wickham, Hester, \& Chang, 2018), \emph{dplyr} (Version
0.8.3; Wickham, François, Henry, \& Müller, 2018), \emph{forcats}
(Version 0.3.0; Wickham, 2018), \emph{ggformula} (Version 0.7.0; D.
Kaplan \& Pruim, 2017), \emph{ggplot2} (Version 3.2.1; Wickham, 2016),
\emph{haven} (Version 2.1.0; Wickham \& Miller, 2019), \emph{irr}
(Version 0.84.1; Gamer, Lemon, \&
\textless{}puspendra.pusp22@gmail.com\textgreater{}, 2012), \emph{knitr}
(Version 1.25; Xie, 2015), \emph{kutils} (Version 1.70; Johnson, Kite,
\& Redmon, 2019), \emph{lattice} (Version 0.20.38; Sarkar, 2008),
\emph{lavaan} (Version 0.6.1; Rosseel, 2012), \emph{lpSolve} (Version
5.6.15; Berkelaar \& others, 2015), \emph{Matrix} (Version 1.2.15; Bates
\& Maechler, 2017), \emph{mosaic} (Version 1.2.0; Pruim, Kaplan, \&
Horton, 2017, 2016), \emph{mosaicData} (Version 0.17.0; Pruim et al.,
2016), \emph{nlme} (Version 3.1.137; Pinheiro, Bates, DebRoy, Sarkar, \&
R Core Team, 2017), \emph{papaja} (Version 0.1.0.9842; Aust \& Barth,
2018), \emph{psych} (Version 1.8.4; Revelle, 2017), \emph{purrr}
(Version 0.3.2; Henry \& Wickham, 2019), \emph{readr} (Version 1.1.1;
Wickham, Hester, \& Francois, 2017), \emph{stringr} (Version 1.4.0;
Wickham, 2019), \emph{tibble} (Version 2.1.3; Müller \& Wickham, 2019),
\emph{tidyr} (Version 1.0.0; Wickham \& Henry, 2019), \emph{tidyverse}
(Version 1.2.1; Wickham, 2017), \emph{usethis} (Wickham \& Bryan, 2018),
and \emph{xtable} (Version 1.8.3; Dahl, Scott, Roosen, Magnusson, \&
Swinton, 2019) for all our analyses.

@@

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential data parter. While we think
that hetereosexual women are indeed able to do this with other women as
their target---indeed, there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---they might be more
apt to activate social comparison processes (Festinger, 1954) than women
who are sexually attracted to other women (non-heterosexual women). This
differential psychological process between women with differing
sexualities might have served to dampen our ability to detect
relationships, adding variability.

>>>>

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for women within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between such valuable phenomena and their effects on authenticity within
interactions between female pairs. Women were brought into the
laboratory and interacted in same-sex dyads. Dyadic analysis was
utilized to detect whether partners' objectification of each other
affected state self-objectification, and the resulting feelings of
comfort and authenticity during the interaction. After the interaction,
participants completed a questionnaire which measured many constructs
including cognitive performance, career aspirations, and relationship
agency. Results revealed no significant relationship between
self-objectification and authenticity. Further, although there were
significantly negative effects on career aspirations and relationship
agency resulting from a lack of relationship authenticity, there was no
evidence that this is due to feelings of sexual objectification. The
significant partner effect of objectification on actor
self-objectification suggests that women being objectified by other
women still results in feelings of self-objectification, and such
research has powerful implications for the ways that women interact in
both sexual and non-sexual settings. AUTHOR NOTE: mention Clark trigram
coders Hannah et al.

@@

Roberts and Fredrickson (1997) Objectification Theory, suggests that, in
addition to being steeped in a culture that objectifies women, women are
objectified in actual interpersonal encounters. The negative effects of
this interpersonal objectification might be the strongest when it is a
man doing the objectifying (Gervais, Vescio, \& Allen, 2011), but what
are the consequences when a woman objectifies another woman in an
interaction? Psychological researchers studying the sexual
objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
self-objectification in the target as a proximal consequence (Garcia,
Earnshaw, \& Quinn, 2016, Strelan and Pagoudis (2018), Riemer, Sáez,
Brock, and Gervais (2020)). For example, in a face-to-face mixed-gender
dyadic study, Garcia et al. (2016) found that men's reported
objectification of their female interaction partner was associated with
increased self-objectification reported by their partner. They did not
include same-gender, woman-woman interacting pairs. Gervais, Sáez,
Riemer, and Klein (2020) have reviewed the literature on interpersonal
self-objectification and has organized the process in a theoretical
model called the Social Interaction Model of Objectification (SIMO), but
this model is currently only relevant for mixed-gender encounters.
Although there is ample evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), there are currently no
studies investigating the process of interpersonal objectification in
same-gender interactions. Does objectification by a female interaction
partner have the same downstream negative consequences as being
objectified by a male interaction partner? The current dyadic study
addresses this omission using interacting pairs where both members
identify as women.

@@

Self-objectification is a psychological process that provides a
cognitive mechanism to translate the experiences of interpersonal
objectification (Loughnan et al., 2015a) to negative mental health
outcomes (e.g., anxiety, self-esteem, and cognitive performance)
(Calogero, Tantleff-Dunn, \& Thompson, 2011; Moradi \& Huang, 2008).
Calogero et al. (2011) proposes that self-objectification can be
conceptualized as a learned trait, or \emph{trait} self-objectification
(TSO). Furthermore, it can also be elicited momentarily, through the
media, for example, with sexualized images in movies and magazines
(Morry \& Staska, 2001), or when wearing sexualizing clothing (B. L.
Fredrickson, Roberts, Noll, Quinn, \& Twenge, 1998). This momentary
self-objectification is referred to as \emph{state} self-objectification
(Bernard, Legrand, \& Klein, 2018; Calogero et al., 2011; Moradi \&
Huang, 2008). Being objectified by another person may interact TSO to
effect experiences of feeling like a body, or state self-objectification
(SSO; Garcia et al., 2016).

@@

Harsey and Zurbriggen (2020) found that self-objectification was related
to the objectification of women to a similar degree for men and women
participants (although men scored higher than women on a measure of
belief that objectification of women is natural for men). Additionally,
Strelan and Hargreaves (2005) found that the more women self-objectify,
the more they objectify other women. Could the process work in the other
direction as well, that is, could objectification by a woman produce
self-objectification, in a similar manner to mixed-gender interactions?

{[}Asha comment on next paragraph: Hill and Fischer (2008) assessed
women's experiences of objectification from men separately from women's
experiences of objectification from women, and found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well, although this process was not comfirmed
to occur during an actual interpersonal interaction

Could this sentence be written like:

Though, when Hill and Fischer(2008) assessed women's experiences of
objectification from men independently from that of women, they found
that women are likely to be socialized to not only to see themselves s
objects, but potentially see other women as objects as well. This
process ,however, was not confirmed to occur during an interpersonal
interaction.{]}

Past research has focused on men's objectifying behaviors, with little
attention paid to the role that women may also play in objectifying
other women. Hill and Fischer (2008) assessed women's experiences of
objectification from men separately from women's experiences of
objectification from women and found that women may be socialized not
only to see themselves as objects but perhaps to see other women as
objects as well, although this process was not comfirmed to occur during
an actual interpersonal interaction. Loughnan et al. (2015a) found that
women objectify other women to a greater extent than they objectify men.
Women have also been found to objectify (dehumanized) sexualized targets
presented as images (Puvia \& Vaes, 2013), but very little is known
about interpersonal sexual objectification among women and what the
proximal effects of this objectification might be. We do know that
objectification has more adverse consequences for women than men
(Gervais et al., 2011; Moradi \& Huang, 2008; Saguy, Quinn, F Dovidio,
\& Pratto, 2010), however, if women objectify other women, perhaps even
to a stronger degree than they objectify men, this higher level of
objectification might be a detrimental combination.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), and often times feel \enquote{looked
at} within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and will more than likely internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Moreover, perhaps the most adverse effect is that it socializes
girls and women to treat themselves as objects to be looked at and
evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972; B. L.
Fredrickson et al., 1998). Gervais et al. (2020) recently reviewed the
research on interpersonal objectification and organized our current
theoretical understanding of this process in the Social Interaction
Model of Objectification (SIMO). What is clear from Gervais et al.
(2020) is that we know little about women-on-women interpersonal
objectification. Could there also be negative consequences of this type
of interpersonal objectification? Perhaps women could be amplifying each
others' state self-objectification as they get ready to go out for the
night {[}EDIT??{]}, \emph{before} the theoretical encounters with men
organized in the SIMO occur. In this way, women's rituals of getting
ready together might place even women with low TSO in a higher SSO
group. {[}Kat: This has me thinking about the difference between
objectification in a romantic setting (which was the setting created in
the initial experiment) vs objectification in a platonic setting/between
friends. Might be useful to discuss this difference, or just have a
sentence here that clarifies objectification can occur within non-sexual
settings{]} {[}Asha: I think this is a great way to illustrate the
potential experience but when I read it I was kind of like
\enquote{wait, when did we talk about women going out?}{]}

Interpersonal objectification can also occur within the context of
romantic relationships, in addition to interactions with strangers.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, suggesting that women
can sometimes have positive outcomes from appearance valuation within
the context of an established romantic relationship. There is also
evidence that they experience less SSO after appearance comments within
romantic relationships (Meltzer, 2020), but less is known about first
dates. An important caveat to this literature is that it is only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) {[}CITE PWQ paper I
reviewed, Yoder editor{]}. Both romantic relationships and stranger
encounters have been studied in the context of objectification, but
neither of these literatures have investigated specifically same-gender
relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? On the one hand, there is ample
research demonstrating that the male gaze has a particularly detrimental
effect would suggest no (Gay \& Castano, 2010; Gervais et al., 2011;
Saguy et al., 2010, Calogero (2004); Yilmaz \& Bozo, 2019), but on the
other hand, women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013). When they do it, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

There is some evidence that reduced authenticity is a consequence of
self-objectification in the moment {[}Garcia et al. (2016); again CITE
PWQ paper I reviewed, Yoder editor{]}. This link has been justified by
the literature on stigmatized-stigmatizer interactions (M. R. Hebl \&
Dovidio, 2005; Richeson \& Shelton, 2003; Shelton, Richeson, \&
Salvatore, 2005), viewing the experience of being objectified in an
interaction as an identity threat situation (Miles-McLean et al., 2015;
Nadal \& Haynes, 2012). Further, empirical evidence reveals that
objectification manifests through inauthenticity in romantic
relationships (Brunell et al., 2010), adverse attitudes in regard to
career aspirations, and a decrease in concentration and impairment in
female cognitive performance (Kahalon, Shnabel, \& Becker, 2018; D. M.
Quinn, Chaudoir, \& Kallen, 2011). When a woman is objectified by a man,
and subsequently experiences self-objectification, the intergroup nature
of this encounter might trigger identity threat. Thus, perhaps there are
fewer negative consequences when a woman is objectified by another
woman, specifically, woman-woman interpersonal objectification processes
might diverge from mix-gender interpersonal objectification processes
precisely because they are not intergroup interactions (at least with
respect to gender identity). A woman objectified by another woman may
not be having the same negative consequences that cascade from
situations that trigger group-based identity threat (Deaux \& Major,
1987; Dovidio, Hebl, Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio,
2005), but feeling like a body, rather than a full human, in any
interaction (intergroup or ingroup) may be enough to reduce women's
feelings of authenticity and social competence, regardless of the gender
of the objectifyer {[}Tolman, Impett, Tracy, and Michael (2006); PWQ
paper I reviewed, Yoder editor{]}.

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected, in line
with Garcia et al. (2016), that TSO would moderate this relationship,
amplifying the positive assocaition between being objectified and SSO
for women higher in TSO. We hypothesized that SSO would, in turn, lead
to feelings of inauthenticity (although there is ample support against
this prediction from the identity threat within intergroup interactions
literature; Rollero, 2016). Further, we hypothesized that these feelings
of inauthencity would then be associated with reduced feelings of agency
in romantic relationships, reduced career aspiration, and reduced
cognitive performance. In summary, we expected to find a positive
relationship between other-objectification by one's partner and state
self-objectification. We also expected to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would predict the TSO to other-objectification
link, further theorezing that this relationships is mediated by the
woman's own state self-objectification (SSO). This limited, but extant,
set of literature on woman-woman objectification points to a possible
alternative model from the model hypothesized in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

@@

Except for the instructions given to participants, the procedure used
was identical to that in Garcia et al. (2016). In brief, that
methodology is that each participant arrived at the laboratory and were
then led into separate cubicles to prevent any communication between the
participants before the interaction. In addition, each participant was
screened for prior acquaintance to confirm that they had not met prior
to the study. They were asked to sign the consent form to participate,
and the study was described as follows: \enquote{This is a study looking
at how students form different types of relationships at college.} A
prompt on the computer screen told the participants that they were
assigned to the \enquote{College Relationships} condition and gave the
following instructions:

\begin{quote}
There are many types of relationships people form in college. During the
interaction, please think about your partner's potential as a romantic
partner. \textbf{Even if they are not the gender you are attracted to},
you can still judge their potential as a romantic partner. After the
interaction you will be asked to evaluate how dateable your partner is.
In other words, we would like to know if you think someone would date
your interaction partner. Also, your interaction partner will be
evaluating you in the same manner.
\end{quote}

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationship, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
women are indeed able to evaluate other women's potential as romantic
partners {[}i.e., their sexual attractiveness; CITE{]}---indeed, women
may be unfortunately quite used to thinking about their own
\enquote{datability,} and we suspect this this habitual thought pattern
will translate to their thoughts about other women.

After recieving the \enquote{College Relationships} prompt, the two
participants were then brought into a larger interaction room where they
sat on stools prearranged to be approximately 1 meter apart. The
experimenter instructed the participants to \enquote{get to know each
other} for 10 minutes and then left the room. After 10 minutes, the
experimenter came back into the room and stopped the interaction. The
participants then went back to their individual cubicles and completed a
set of post-interaction measures. Participants were then thanked for
their participation and debriefed. More detail on this methodology can
be found in Garcia et al. (2016).

@@

Due to these similarities across samples in regard to correlation
patterns between significant variables within this study, the two
datasets were combined. These participants were mostly first-year
college students, with an average age of 18.85 (SD = 1.04). The sample
was 48.44\% White/European American, 9.38\% Black/African-American,
28.12\% Asian/Pacific Islander, 9.38\% Latinx, and 4.69\% mixed-race.
There were 8 White/White pairs and 4 same race racial minority pairs,
for a total of 12 same-race pairs. The remaining 20 were mixed race
pairs, of which 15 were White/racial minority pairings and 5 were
cross-racial minority group pairs. 64.06\% of the sample identified as
heterosexual, and 25\% identified as gay, lesbian or bisexual.

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner (Garcia et al., 2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were rated on a
scale from 1 (not at all) to 7 (constantly). Objectification was
measured by getting the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

@@

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much alike (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{`Do you think your partner
was authentic during your interaction?}' and \enquote{`Were you
authentic during your interaction?}' These questions were ranked on a
scale from 1 (not at all) to 7 (very much). These were combined to form
the authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) =
0.91 for Sample 2).

@@

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{tidyverse} (Version 1.2.1; Wickham, 2017), \emph{psych} (Version
1.8.4; Revelle, 2017), \emph{nlme} (Version 3.1.137; Pinheiro, Bates,
DebRoy, Sarkar, \& R Core Team, 2017) for our analyses. Further, we used
the R-packages \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
\emph{apaTables} (Version 2.0.5; Stanley, 2018), with \emph{knitr}
(Version 1.25; Xie, 2015) to create a fully reproducible manuscript in
\emph{rmarkdown} (Version 1.15; Xie, Allaire, \& Grolemund, 2018).

@@

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential dating/romantic partner. We
think that hetereosexual women are able to do this with other women as
their target---and there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---they may be even
more apt to activate social comparison processes (Festinger, 1954) than
women who are sexually attracted to other women (non-heterosexual
women). This differential psychological process between women with
differing sexualities might have served to dampen our ability to detect
relationships, adding variability. {[}Kat: The only other thing I would
add re: sexuality limitations are that we didn't collect/use sexual
orientation, so we couldn't analyze results by that at all/our samples
might have differed there.{]}
